Scottish Child Poverty Briefing Plots

### 2024-06-05 minos/outcomes/ChildPovReduction_Combined.Rmd
Updating the ChildPovReduction plotting script with num_priority_groups plots and adding a new script with both relative and absolute interventions together

```{r "SF12_change_simple_functions", include=FALSE}
SF12_change_simple <- function(summary_group, scen.list) {

  plot_data <- data %>%
    select(run_id, SF_12, mean_cost, year, scen) %>%
    pivot_wider(names_from = scen,
                values_from = c(SF_12, mean_cost),
                names_sep = '.') %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    pivot_longer(cols = starts_with("diff") | starts_with("mean_cost"), 
                 names_to = c(".value", "scen"), 
                 names_sep = "\\.") %>%
    filter(scen != 'baseline')

  plot_data_confint <- plot_data %>%
    group_by(year) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF_12),
              margin = qt(0.975, df = n - 1) * (sd(diff_SF_12) / sqrt(n)),
              lower = mean_SF12_change - margin,
              upper = mean_SF12_change + margin)

  plot_data_confint2 <- plot_data %>%
    group_by(year) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF_12),
              margin = qnorm(0.975) * (sd(diff_SF_12) / sqrt(n)),
              lower = mean_SF12_change - margin,
              upper = mean_SF12_change + margin)

  p1 <- ggplot(plot_data, aes(x = year, y = diff_SF_12)) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted') +
    geom_smooth() +
    labs(title = tit) +
    xlab('Year') +
    ylab('Change in SF-12')

  p2 <- ggplot(plot_data_confint, aes(x = year, y = mean_SF12_change)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), linetype = 'dotted', alpha = 0.3) +
    geom_line(linetype = 'dashed') +
    geom_point() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted', alpha = 0.5) +
    labs(title = tit) +
    xlab('Year') +
    ylab('Change in SF-12')

  p3 <- ggplot(plot_data_confint2, aes(x = year, y = mean_SF12_change)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), linetype = 'dotted', alpha = 0.3) +
    geom_line(linetype = 'dashed') +
    geom_point() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted', alpha = 0.5) +
    labs(title = tit) +
    xlab('Year') +
    ylab('Change in SF-12')

  print(p1)
  print(p2)
  #print(p3)
}

SF12_change_percentage <- function(summary_group, scen.list) {

@@

  p1 <- ggplot(plot_data, aes(x = year, y = percent_change_SF_12)) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted') +
    geom_smooth() +
    labs(title = tit) +
    xlab('Year') +
    ylab('Percentage Change in SF-12')

  p2 <- ggplot(plot_data_confint, aes(x = year, y = mean_SF12_change)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), linetype = 'dotted', alpha = 0.3) +
    geom_line(linetype = 'dashed') +
    geom_point() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted', alpha = 0.5) +
    labs(title = tit) +
    xlab('Year') +
    ylab('Percentage Change in SF-12')

  print(p1)
  print(p2)
}

```{r "SF12_change_income_quint", include=FALSE}
SF12_change_income_quint <- function(summary_group, scen.list) {

  plot_data <- data %>%
    select(run_id, income_quintile, SF_12, mean_cost, year, scen) %>%
    pivot_wider(names_from = scen,
                values_from = c(SF_12, mean_cost),
                names_sep = '.') %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    pivot_longer(cols = starts_with("diff") | starts_with("mean_cost"),
                 names_to = c(".value", "scen"),
                 names_sep = "\\.") %>%
    filter(scen != 'baseline') %>%
    select(-scen)

  plot_data_confint <- plot_data %>%
    group_by(year, income_quintile) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF_12),
              margin = qt(0.975, df = n - 1) * (sd(diff_SF_12) / sqrt(n)),
              lower = mean_SF12_change - margin,
              upper = mean_SF12_change + margin)

  plot_data_confint2 <- plot_data %>%
    group_by(year, income_quintile) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF_12),
              margin = qnorm(0.975) * (sd(diff_SF_12) / sqrt(n)),
              lower = mean_SF12_change - margin,
              upper = mean_SF12_change + margin)

  p1 <- ggplot(plot_data, aes(x = year, y = diff_SF_12, group = income_quintile, colour = income_quintile, fill = income_quintile)) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted') +
    geom_smooth() +
    labs(title = tit, subtitle = 'Income Quintile') +
    xlab('Year') +
    ylab('Change in SF-12')

  p2 <- ggplot(plot_data_confint, aes(x = year, y = mean_SF12_change, group = income_quintile, colour = income_quintile, fill = income_quintile)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), linetype = 'dotted', alpha = 0.3) +
    geom_line(linetype = 'dashed') +
    geom_point() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted', alpha = 0.5) +
    labs(title = tit, subtitle = 'Income Quintile') +
    xlab('Year') +
    ylab('Change in SF-12')

  p3 <- ggplot(plot_data_confint2, aes(x = year, y = mean_SF12_change, group = income_quintile, colour = income_quintile, fill = income_quintile)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), linetype = 'dotted', alpha = 0.3) +
    geom_line(linetype = 'dashed') +
    geom_point() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted', alpha = 0.5) +
    labs(title = tit, subtitle = 'Income Quintile') +
    xlab('Year') +
    ylab('Change in SF-12')

  print(p1)
  print(p2)
}

SF12_change_percentage_income_quint <- function(summary_group, scen.list) {

@@

  print(p1)
  print(p2)
}

```{r "SF12_change_num_priority_groups", include=FALSE}
SF12_change_num_priority_groups <- function(scen.list) {

@@

  data <- do.call(rbind, generate_df_list(summary_group, scen.list.test))

  plot_data <- data %>%
    select(run_id, num_priority_groups, SF_12, mean_cost, year, scen) %>%
    pivot_wider(names_from = scen,
                values_from = c(SF_12, mean_cost),
                names_sep = '.') %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    pivot_longer(cols = starts_with("diff") | starts_with("mean_cost"),
                 names_to = c(".value", "scen"),
                 names_sep = "\\.") %>%
    filter(scen != 'baseline') %>%
    select(-scen)

  plot_data_confint <- plot_data %>%
    group_by(year, num_priority_groups) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF_12),
              margin = qt(0.975, df = n - 1) * (sd(diff_SF_12) / sqrt(n)),
              lower = mean_SF12_change - margin,
              upper = mean_SF12_change + margin)

  plot_data_confint2 <- plot_data %>%
    group_by(year, num_priority_groups) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF_12),
              margin = qnorm(0.975) * (sd(diff_SF_12) / sqrt(n)),
              lower = mean_SF12_change - margin,
              upper = mean_SF12_change + margin)

  p1 <- ggplot(plot_data, aes(x = year, y = diff_SF_12, group = num_priority_groups, colour = num_priority_groups, fill = num_priority_groups)) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted') +
    geom_smooth() +
    labs(title = tit) +
    xlab('Year') +
    ylab('Change in SF-12')

  print(p1)
  print(p2)
}

SF12_change_percentage_num_priority_groups <- function(scen.list) {

  data <- do.call(rbind, generate_df_list(summary_group, scen.list.test))

@@

  p1 <- ggplot(plot_data, aes(x = year, y = diff_SF_12, group = num_priority_groups, colour = num_priority_groups, fill = num_priority_groups)) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted') +
    geom_smooth() +
    labs(title = tit) +
    xlab('Year') +
    ylab('Change in SF-12')

@@

```{r}
#scen.list.test <- c('baseline', 'ChildPovertyReductionRELATIVE_2')
scen.list.test <- c('baseline', 'ChildPovertyReductionABSOLUTE')
whole_pop_data <- do.call(rbind, generate_df_list('whole_pop', scen.list.test))
```

# Quick Data Test

```{r}
ggplot(whole_pop_data, aes(x = year, y = SF_12, group = scen, colour = scen, fill = scen)) +
  geom_smooth()
```

Check if SF12 results in one year (across 100 runs) are normally distributed.

```{r}
whole_pop_2030 <- whole_pop_data %>%
  filter(year == 2030)

hist(whole_pop_2030$SF_12)
qqnorm(whole_pop_2030$SF_12)
qqline(whole_pop_2030$SF_12)

shapiro.test(whole_pop_2030$SF_12)

# Results look normally distributed^
```

# Whole Population

@@

```{r}
SF12_change_simple(summary_group = 'whole_pop',
                   scen.list = scen.list.test)

SF12_change_percentage(summary_group = 'whole_pop',
                       scen.list = scen.list.test)
```

```{r}
SF12_change_income_quint(summary_group = 'whole_pop',
                         scen.list = scen.list.test)

SF12_change_percentage_income_quint(summary_group = 'whole_pop',
                                    scen.list = scen.list.test)
```

# Families

@@

```{r}
SF12_change_simple(summary_group = 'families',
                   scen.list = scen.list.test)

SF12_change_percentage(summary_group = 'families',
                       scen.list = scen.list.test)
```

```{r}
SF12_change_income_quint(summary_group = 'families',
                         scen.list = scen.list.test)

SF12_change_percentage_income_quint(summary_group = 'families',
                                    scen.list = scen.list.test)
```

# Priority Subgroups

@@

```{r}
SF12_change_simple(summary_group = 'priority_any',
                         scen.list = scen.list.test)

SF12_change_percentage(summary_group = 'priority_any',
                                    scen.list = scen.list.test)
```

## Number of Priority Subgroups

@@

# TESTING

# Differential Impacts of UK vs SG

# BONUS - Treated Pop

>>>>

```{r "SF12_change_simple_functions", include=FALSE}
SF12_change_percentage <- function(summary_group, scen.list) {

@@

  p1 <- ggplot(plot_data, aes(x = year, y = percent_change_SF_12)) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted') +
    geom_smooth() +
    labs(title = tit, subtitle = 'GAM Smoothed') +
    xlab('Year') +
    ylab('Percentage Change in SF-12')

  p2 <- ggplot(plot_data_confint, aes(x = year, y = mean_SF12_change)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), linetype = 'dotted', alpha = 0.3) +
    geom_line(linetype = 'dashed') +
    geom_point() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted', alpha = 0.5) +
    labs(title = tit, subtitle = '95% Confidence Intervals') +
    ylab('Percentage Change in SF-12')

```{r "SF12_change_income_quint", include=FALSE}
SF12_change_percentage_income_quint <- function(summary_group, scen.list) {

@@

```{r "SF12_change_num_priority_groups", include=FALSE}
SF12_change_percentage_num_priority_groups <- function(scen.list) {

@@

@@

  p1 <- ggplot(plot_data, aes(x = year, y = percent_change_SF_12, group = num_priority_groups, colour = num_priority_groups, fill = num_priority_groups)) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted') +
    geom_smooth() +
    labs(title = tit) +
    xlab('Year') +
    ylab('Change in SF-12')

@@

```{r "Data", include=FALSE}
scen.list.rel <- c('baseline', 'ChildPovertyReductionRELATIVE_2')
scen.list.abs <- c('baseline', 'ChildPovertyReductionABSOLUTE')
```

# RELATIVE POVERTY INTERVENTION

Reducing the proportion of children in poverty to 10% or lower by 2030.

## Whole Population

@@

```{r}
SF12_change_percentage(summary_group = 'whole_pop',
                   scen.list = scen.list.rel)
```

```{r}
SF12_change_percentage_income_quint(summary_group = 'whole_pop',
                                    scen.list = scen.list.rel)
```

## Families

@@

```{r}
SF12_change_percentage(summary_group = 'families',
                       scen.list = scen.list.rel)
```

```{r}
SF12_change_percentage_income_quint(summary_group = 'families',
                                    scen.list = scen.list.rel)
```

## Priority Subgroups

@@

```{r}
SF12_change_percentage(summary_group = 'priority_any',
                                    scen.list = scen.list.rel)
```

### Number of Priority Subgroups

Results disaggregated by if household belongs to 0, 1, 2, 3+ priority subgroups.

```{r}
SF12_change_percentage_num_priority_groups(scen.list.rel)
```

## Differential Impacts of UK vs SG

# ABSOLUTE POVERTY INTERVENTION

Reducing the proportion of children in poverty to 10% or lower by 2030.

## Whole Population

```{r}
SF12_change_percentage(summary_group = 'whole_pop',
                       scen.list = scen.list.abs)
```

```{r}
SF12_change_percentage_income_quint(summary_group = 'whole_pop',
                                    scen.list = scen.list.abs)
```

## Families

```{r}
SF12_change_percentage(summary_group = 'families',
                       scen.list = scen.list.abs)
```

```{r}
SF12_change_percentage_income_quint(summary_group = 'families',
                                    scen.list = scen.list.abs)
```

## Priority Subgroups

```{r}
SF12_change_percentage(summary_group = 'priority_any',
                                    scen.list = scen.list.abs)
```

### Number of Priority Subgroups

Results disaggregated by if household belongs to 0, 1, 2, 3+ priority subgroups.

```{r}
SF12_change_percentage_num_priority_groups(scen.list.abs)
```

## Differential Impacts of UK vs SG

@@

### 2024-06-05 minos/outcomes/ChildPovReduction_plots.Rmd
Updating the ChildPovReduction plotting script with num_priority_groups plots and adding a new script with both relative and absolute interventions together

@@

```{r}
#scen.list.test <- c('baseline', 'ChildPovertyReductionRELATIVE_2')
scen.list.test <- c('baseline', 'ChildPovertyReductionABSOLUTE')
whole_pop_data <- do.call(rbind, generate_df_list('whole_pop', scen.list.test))
```

@@

>>>>

  p1 <- ggplot(plot_data, aes(x = year, y = percent_change_SF_12, group = num_priority_groups, colour = num_priority_groups, fill = num_priority_groups)) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted') +
    geom_smooth() +
    labs(title = tit) +
    xlab('Year') +
    ylab('Change in SF-12')

@@

```{r}
scen.list.test <- c('baseline', 'ChildPovertyReductionRELATIVE_2')
#scen.list.test <- c('baseline', 'ChildPovertyReductionABSOLUTE')
whole_pop_data <- do.call(rbind, generate_df_list('whole_pop', scen.list.test))
```

@@

```{r}
SF12_change_num_priority_groups(scen.list.test)
SF12_change_percentage_num_priority_groups(scen.list.test)
```
